Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-01-31 WSSSPE1_report/WSSSPE1_summary_report.tex
adding use-case section
reorganizing

People who have volunteered to work on this report:

The workshop organizers:
\begin{itemize}
\item Daniel S. Katz $\langle$\url{dkatz@nsf.gov}$\rangle$
\item Gabrielle Allen $\langle$\url{gdallen@illinois.edu}$\rangle$
\item Neil Chue Hong $\langle$\url{N.ChueHong@software.ac.uk}$\rangle$
\item Manish Parashar $\langle$\url{parashar@rutgers.edu}$\rangle$
\item David Proctor $\langle$\url{dproctor@nsf.gov}$\rangle$
\end{itemize}

Additional volunteers:
\begin{itemize}
\item Chris Mattmann $\langle$\url{chris.a.mattmann@jpl.nasa.gov}$\rangle$
\item Ketan Maheshwari $\langle$\url{ketancmaheshwari@gmail.com}$\rangle$
\item Marlon Pierce $\langle$\url{marpierc@iu.edu}$\rangle$
\item Colin Venters $\langle$\url{C.Venters@hud.ac.uk}$\rangle$
\item Suresh Marru $\langle$\url{smarru@iu.edu}$\rangle$
\item Lynn Zentner $\langle$\url{lzentner@purdue.edu}$\rangle$
\item Anne Elster $\langle$\url{anne.elster@gmail.com}$\rangle$
\item Shel Swenson $\langle$\url{mdswenso@usc.edu}$\rangle$
\item Andy Ray Terrel $\langle$\url{andy.terrel@gmail.com}$\rangle$
\item Abani Patra $\langle$\url{abani.patra@gmail.com}$\rangle$
\item Nancy Wilkins-Diehr $\langle$\url{wilkinsn@sdsc.edu}$\rangle$
\item James Spencer $\langle$\url{j.spencer@imperial.ac.uk}$\rangle$
\item Frank Loeffler $\langle$\url{knarf@cct.lsu.edu}$\rangle$
\item James Hetherington $\langle$\url{j.hetherington@ucl.ac.uk}$\rangle$
\item Sou-Cheng (Terrya) Choi $\langle$\url{sou.cheng.terrya.choi@gmail.com}$\rangle$
\item James Howison $\langle$\url{james@howison.name}$\rangle$
\item Bruce Berriman $\langle$\url{gbb@ipac.caltech.edu}$\rangle$
\item Hilmar Lapp $\langle$\url{hlapp@nescent.org}$\rangle$
\item Marcus D. Hanwell $\langle$\url{marcus.hanwell@kitware.com}$\rangle$
\item Lucas Nussbaum $\langle$\url{lucas.nussbaum@loria.fr}$\rangle$
\item Matthew Turk $\langle$\url{matthewturk@gmail.com}$\rangle$
\end{itemize}

@@

@@

...

\section{Agenda}
The workshop submissions were classified into various related categories.
Representatives from each submission categories were appointed as penalists. The sessions were kickstarted with keynote talks followed by panel discussions.

@@

what was presented and what was discussed

@@

what was presented and what was discussed

@@

What is sustainable? What are the best practices, can workshops fund meetings to help define and improve best practices in these areas? Software plans from the funding agencies?

@@

what was presented and what was discussed

@@

what was presented and what was discussed

@@

from the misc questions/answers/observations part of the google doc?

@@

@@

pull the discussion together

@@

if needed.

@@

or at least the discussion about them, and the current plans for future events.

@@

\section{Pre-workshop section: Introduction}

Progress in scientific research is dependent on the quality and accessibility of software at all levels and it is now critical to address many challenges related to the development, deployment, and maintenance of reusable software. In addition, it is essential that scientists, researchers, and students are able to learn and adopt software-related skills and methodologies. Established researchers are already acquiring some of these skills, and in particular a specialized class of software developers is emerging in academic environments who are an integral and embedded part of successful research teams. The WSSSPE workshop is intended provide a forum for discussion of the challenges, including both positions and experiences. The short papers and discussion will be archived to provide a basis for continued discussion, and the workshop is intended to feed into the collaborative writing of one or more journal publications.

\section{Pre-workshop section: Submissions}

\section{Pre-workshop section: Peer-Review}

\section{Pre-workshop section: Results}

The contributed papers that were discussed at the workshop follow, listed by area.

\subsection{Developing, Deploying and Supporting Software}

\subsubsection{Development Experiences}

\subsubsection{Deployment, Support, and Maintenance of Existing Software}

%\end{itemize}
%
%\subsubsection{Tools}

\subsubsection{Best Practices, Challenges, and Recommendations}

\subsection{Policy}

\subsubsection{Modeling Sustainability}

\subsubsection{Credit, Citation, Impact}

In addition, the following paper from another area will also be discussed in this area.

\subsubsection{Reproducibility}

\subsubsection{Implementing Policy}

\subsection{Communities}

\subsubsection{Communities}

In addition, the following paper from another area will also be discussed in this area.

\subsubsection{Industry \& Economic Models}

In addition, the following papers from other areas will also be discussed in this area.

\subsubsection{Education \& Training}

In addition, the following papers from other areas will also be discussed in this area.

\section{Pre-workshop section: Conclusions}

>>>>

People who have volunteered to work on this report: (with * for actual contributors - \note{if you contribute, add a * by your name})

The workshop organizers:
\begin{itemize}
\item * Daniel S. Katz $\langle$\url{dkatz@nsf.gov}$\rangle$
\item Gabrielle Allen $\langle$\url{gdallen@illinois.edu}$\rangle$
\item Neil Chue Hong $\langle$\url{N.ChueHong@software.ac.uk}$\rangle$
\item Manish Parashar $\langle$\url{parashar@rutgers.edu}$\rangle$
\item David Proctor $\langle$\url{dproctor@nsf.gov}$\rangle$
\end{itemize}

Additional volunteers:
\begin{itemize}
\item Chris Mattmann $\langle$\url{chris.a.mattmann@jpl.nasa.gov}$\rangle$
\item * Ketan Maheshwari $\langle$\url{ketancmaheshwari@gmail.com}$\rangle$
\item Marlon Pierce $\langle$\url{marpierc@iu.edu}$\rangle$
\item Colin Venters $\langle$\url{C.Venters@hud.ac.uk}$\rangle$
\item Suresh Marru $\langle$\url{smarru@iu.edu}$\rangle$
\item Lynn Zentner $\langle$\url{lzentner@purdue.edu}$\rangle$
\item Anne Elster $\langle$\url{anne.elster@gmail.com}$\rangle$
\item Shel Swenson $\langle$\url{mdswenso@usc.edu}$\rangle$
\item Andy Ray Terrel $\langle$\url{andy.terrel@gmail.com}$\rangle$
\item Abani Patra $\langle$\url{abani.patra@gmail.com}$\rangle$
\item Nancy Wilkins-Diehr $\langle$\url{wilkinsn@sdsc.edu}$\rangle$
\item James Spencer $\langle$\url{j.spencer@imperial.ac.uk}$\rangle$
\item Frank Loeffler $\langle$\url{knarf@cct.lsu.edu}$\rangle$
\item James Hetherington $\langle$\url{j.hetherington@ucl.ac.uk}$\rangle$
\item * Sou-Cheng (Terrya) Choi $\langle$\url{sou.cheng.terrya.choi@gmail.com}$\rangle$
\item James Howison $\langle$\url{james@howison.name}$\rangle$
\item * Bruce Berriman $\langle$\url{gbb@ipac.caltech.edu}$\rangle$
\item Hilmar Lapp $\langle$\url{hlapp@nescent.org}$\rangle$
\item * Marcus D. Hanwell $\langle$\url{marcus.hanwell@kitware.com}$\rangle$
\item Lucas Nussbaum $\langle$\url{lucas.nussbaum@loria.fr}$\rangle$
\item Matthew Turk $\langle$\url{matthewturk@gmail.com}$\rangle$
\end{itemize}

@@

\todo{1 page summary goes here}

@@

Because progress in scientific research is dependent on the quality and accessibility
of software at all levels, it is now critical to address many challenges related to the
development, deployment, and maintenance of reusable software. In addition, it is
essential that scientists, researchers, and students are able to learn and adopt
software-related skills and methodologies. Established researchers are already
acquiring some of these skills, and in particular a specialized class of software
developers is emerging in academic environments who are an integral and embedded
part of successful research teams. The WSSSPE workshop was intended provide a
forum for discussion of the challenges, including both positions and experiences.
The short papers and discussion have been archived to provide a basis for continued
discussion, and the workshop has fed into the collaborative writing of this journal publication.
Additional papers based on extended versions of workshop submissions are expected.
The level of interest in the workshop has led the organizers, working with some of the submitters
and some of the attendees, to plan two additional workshops (\todo{details}) and to turn
the workshop website into a community website that can be used as a focus for further
discussion and progress.

\section{Workshop Process}

WSSSPE1 was organized via peer-review and organization of papers submitted in response to a call for papers.  The call for paper was created by the organizers and peer-review committee.

\section{Agenda}

The workshop consisted of 2 keynote presentations and three panel/discussion sessions.  The panels 
were organized based on a classification of the workshop submissions into three categories, following 
the themes of the call for papers as modified by the areas of the submissions.  Representatives from 
each submission categories were appointed as panelists, to discuss the papers in that category.

@@

\todo{section should include what was presented and what was discussed}

@@

\todo{section should include what was presented and what was discussed}

\subsection{Papers}

\subsubsection*{Development Experiences}

\subsubsection*{Deployment, Support, and Maintenance of Existing Software}

\subsubsection*{Best Practices, Challenges, and Recommendations}

@@

What is sustainable? What are the best practices, can workshops fund meetings to help define and improve best practices in these areas? Software plans from the funding agencies? \todo{mention as cross-cutting, since this came up in other parts of the workshop as well?}

@@

\todo{section should include what was presented and what was discussed}

\subsection{Papers}

\subsubsection*{Modeling Sustainability}

\subsubsection*{Credit, Citation, Impact}

In addition, the following paper from another area were also discussed in this area.

\subsubsection*{Reproducibility}

\subsubsection*{Implementing Policy}

@@

\todo{section should include what was presented and what was discussed}

\subsection{Papers}

\subsubsection*{Communities}

In addition, the following paper from another area was also discussed in this area.

\subsubsection*{Industry \& Economic Models}

In addition, the following papers from other areas were also discussed in this area.

\subsubsection*{Education \& Training}

In addition, the following papers from other areas were also discussed in this area.

@@

\note{from the misc questions/answers/observations part of the google doc?}

@@

\subsection{Defining sustainability}

\subsection{Licensing}

\note{not sure if this was really cross-cutting...}

@@

\todo{pull the discussion together}

\todo{add some analysis of the deficiencies and difficulties that are present in different fields, and those that are common?}

@@

\note{if needed.}

@@

\note{or at least the discussion about them, and the current plans for future events.}

\note{conclusions from pre-workshop paper follows}

@@
